---
<<<<<<< HEAD
title: "A Short Introduction to the formulaic Package"
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{A Short Introduction to the formulaic Package}
output:
  knitr:::html_vignette
---

=======
title: "formulaic"
output: rmarkdown::html_vignette
author: " "
date: "`r Sys.time()`"
vignette: >
  %\VignetteIndexEntry{formulaic}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---
>>>>>>> origin
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  message = FALSE,
  warning = FALSE,
  error = FALSE,
  comment = "#>"
)
<<<<<<< HEAD

#tools::compactPDF(paths = "../vignettes/",
#                  qpdf = Sys.which(Sys.getenv("R_QPDF", "qpdf")),
#                  gs_quality = 'ebook')
```


=======
tools::compactPDF(paths = "../vignettes/",
                  qpdf = Sys.which(Sys.getenv("R_QPDF", "qpdf")),
                  gs_quality = 'ebook')
```

>>>>>>> origin
```{r setup}
library(formulaic)
library(data.table)
data(snack.dat, package = "formulaic")
```

```{r constant}
outcome.name.awareness <- "Awareness"
outcome.name.satisfaction <- "Satisfaction"
<<<<<<< HEAD
input.names <- c("Age", "Gender", "Income", "Region", "Persona", "Typo")
input.patterns <- c("User", "BP_", outcome.name.awareness, outcome.name.satisfaction)
=======

input.names <- c("Age", "Gender", "Income", "Region", "Persona", "Typo")
input.patterns <- c("User", "BP_", outcome.name.awareness, outcome.name.satisfaction)

>>>>>>> origin
reduce <- FALSE
max.input.categories = 20
max.outcome.categories.to.search = 4
order.as = "as.specified" 
include.backtick = "as.needed"
format.as = "formula"
```

## Data : snack.dat
To explore the basic function of formulaic, we’ll use snack.dat, a randomly generated data. This dataset contains 23000 observation and 23 columns. This data contains information from the (fictionalized) marketing survey’s data.
<<<<<<< HEAD

=======
>>>>>>> origin
```{r snack.dat}
dim(snack.dat)
names(snack.dat)
```

<<<<<<< HEAD
## Create Formula Examples

the addback tick examples 

```{r add.backtick}
formulaic:::add.backtick(x = names(snack.dat),include.backtick = 'all')

formulaic:::add.backtick(x = names(snack.dat),include.backtick = 'as.needed')
```

```{r create.formula examples 1}
form.1 <-
  create.formula(
    outcome.name = outcome.name.awareness,
    input.names = NA,
    input.patterns = NA,
    dat = snack.dat
  )

form.1$formula
form.1$inclusion.table
```

```{r create.formula examples 2}
form.2 <-
  create.formula(
    outcome.name = outcome.name.awareness,
    input.names = ".",
    input.patterns = NA,
    dat = snack.dat
  )

form.2$formula
form.2$inclusion.table
```


```{r create.formula examples 3}
form.3 <-
  create.formula(
    outcome.name = outcome.name.awareness,
    input.names = input.names,
    input.patterns = NA,
    dat = NA,
    reduce = reduce,
    max.input.categories = max.input.categories,
    max.outcome.categories.to.search = max.outcome.categories.to.search,
    order.as = order.as,
    include.backtick = include.backtick,
    format.as = "character"
  )

form.3$formula
form.3$inclusion.table
```


```{r create.formula examples 4}
form.4 <-
  create.formula(
    outcome.name = outcome.name.awareness,
    input.names = input.names,
    input.patterns = input.patterns,
    dat = snack.dat,
    reduce = reduce,
    max.input.categories = max.input.categories,
    max.outcome.categories.to.search = max.outcome.categories.to.search,
    order.as = order.as,
    include.backtick = include.backtick,
    format.as = format.as
  )

form.4$formula
form.4$inclusion.table
```


```{r create.formula examples 5}
form.5 <-
  create.formula(
    outcome.name = outcome.name.awareness,
    input.names = input.names,
    input.patterns = input.patterns,
    dat = snack.dat,
    reduce = TRUE,
    max.input.categories = max.input.categories,
    max.outcome.categories.to.search = max.outcome.categories.to.search,
    order.as = order.as,
    include.backtick = include.backtick,
    format.as = format.as
  )

form.5$formula
form.5$inclusion.table
```


```{r create.formula examples 6}
form.6 <-
  create.formula(
    outcome.name = outcome.name.satisfaction,
    input.names = input.names,
    input.patterns = input.patterns,
    dat = snack.dat,
    reduce = TRUE,
    max.input.categories = max.input.categories,
    max.outcome.categories.to.search = max.outcome.categories.to.search,
    order.as = order.as,
    include.backtick = include.backtick,
    format.as = format.as
  )

form.6$formula
form.6$inclusion.table
```



# Reduce Formula Example

```{r reduce.existing.formula examples 7}
form.7 <-
  reduce.existing.formula(
    the.initial.formula = form.3$formula,
    dat = snack.dat,
    format.as = "character"
  )

form.7$formula
form.7$inclusion.table
```

```{r create.formula examples 8}
form.8 <-
  reduce.existing.formula(
    the.initial.formula = form.4$formula,
    dat = snack.dat,
    format.as = "formula"
  )

form.8$formula
form.8$inclusion.table
```

=======
## add.backtick
```{r add.backtick}
formulaic:::add.backtick(x = names(snack.dat))

```

```{r create.formula examples}
form.1 <- create.formula(outcome.name = outcome.name.awareness, input.names = NA, input.patterns = NA, dat = snack.dat)

form.2 <- create.formula(outcome.name = outcome.name.awareness, input.names = ".", input.patterns = NA, dat = snack.dat)


form.3 <- create.formula(outcome.name = outcome.name.awareness, input.names = input.names, input.patterns = NA, dat = NA, reduce = reduce, max.input.categories = max.input.categories, max.outcome.categories.to.search = max.outcome.categories.to.search, order.as = order.as, include.backtick = include.backtick, format.as = "character")

form.4 <- create.formula(outcome.name = outcome.name.awareness, input.names = input.names, input.patterns = input.patterns, dat = snack.dat, reduce = reduce, max.input.categories = max.input.categories, max.outcome.categories.to.search = max.outcome.categories.to.search, order.as = order.as, include.backtick = include.backtick, format.as = format.as)


form.5 <- create.formula(outcome.name = outcome.name.awareness, input.names = input.names, input.patterns = input.patterns, dat = snack.dat, reduce = TRUE, max.input.categories = max.input.categories, max.outcome.categories.to.search = max.outcome.categories.to.search, order.as = order.as, include.backtick = include.backtick, format.as = format.as)

form.6 <- create.formula(outcome.name = outcome.name.satisfaction, input.names = input.names, input.patterns = input.patterns, dat = snack.dat, reduce = TRUE, max.input.categories = max.input.categories, max.outcome.categories.to.search = max.outcome.categories.to.search, order.as = order.as, include.backtick = include.backtick, format.as = format.as)

```

```{r reduce.existing.formula examples}
form.7 <- reduce.existing.formula(the.initial.formula = form.3$formula, dat = snack.dat, format.as = "character")

form.8 <- reduce.existing.formula(the.initial.formula = form.4$formula, dat = snack.dat, format.as = "formula")

```
>>>>>>> origin
