---
title: "Introduction to formualic"
output: rmarkdown::html_vignette
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{Introduction to formualic}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(formulaic)
library(data.table)
library(knitr)
data("snack.dat", package = "formulaic")
```

```{r constant}
outcome.name.awareness <- "Awareness"
outcome.name.satisfaction <- "Satisfaction"
duplicated.inputs <- rep.int(x = "Age", times = 2)
duplicated.inputs2 <- c(rep.int(x = "Age", times = 2), "Income")
duplicated.interactions <- list(c("Age", "Income"), c("Age", "Income"))
input.names <- c("Age", "Gender", "Income", "Region", "Persona", "Typo")
input.patterns <- c("User", "BP_", outcome.name.awareness, outcome.name.satisfaction)
interactions <- list(c("Age", "Gender"), c("Age", "Income"), c("Age", "Gender", "Income"), c("Gender", "Inco"), c("Age", "Reg ion"))
variables.to.exclude = c("Age", "Gender")
reduce <- FALSE
max.input.categories = 20
max.outcome.categories.to.search = 4
order.as = "as.specified"
include.backtick = "as.needed"
format.as = "formula"
force.main.effects <- TRUE
```

## Data : snack.dat

The snack.dat was provided as an example on which to explore the basic functions of the formulaic package. Formatted as a data.table object, snack.dat contains 23000 observation and 23 columns. These data contain simulated information from a fictionalized marketing survey.  In this survey, a progression of questions are asked about the respondents' awareness, consideration, consumption, satisfaction with, and advocacy for different brands of snack foods.  Questions downstream of awareness, then consideration, and then consumption are only asked for those respondents who responded affirmatively to the previous question.  Otherwise the values are missing.  Brand Perception questions are rated on a scale from 0 to 10 and indicated with a name starting with the prefix BP.


```{r snack.dat}
dim(snack.dat)
names(snack.dat)
head(snack.dat)
```

## add.backtick Examples


Add backtick is subsidiary function to create.formula that add backticks to the names of the variables.  Formula objects include the names of different variables within a data.frame.  When these names contain a space, the name must be encapsulated in backticks to ensure proper formatting.  For instance, if there are three variables called y, x1, and User ID, then a formula formatted as y ~ x1 + User ID will generate errors due to the space in User ID.  Instead, this formula can be properly formatted as y ~ x1 + `User ID`.  Meanwhile, it is also acceptable to add backticks to the other names, such as `y` ~ `x1` + `User ID`, but this is not a necessary step.  As a default, the include.backtick is set to 'as.needed', which indicates that the function will only add backticks to the variables that require them. The user has the freedom to change the option to 'all'. Yet, it is only compatible when format.as != "formula", in which case a character object is returned.  In particular, a formula object will automatically remove unnecessary backticks.


In the snack.dat data, User ID is the only variable that is affected by the function when the included.backtick is set as 'as.needed', while every variable has backticks when it is set as 'all'. 


```{r add.backtick}
formulaic:::add.backtick(x = names(snack.dat),include.backtick = 'all')
formulaic:::add.backtick(x = names(snack.dat),include.backtick = 'as.needed')
```

## create.formula examples

The create.formula function was designed to automatically generate formulas from user-specified inputs and outputs.  The range of inputs may include directly specified variables, patterns to search within the names of an associate data.frame, a list of interactions, and a vector of variables to directly exclude from consideration.  The method also provides a range of quality checks that can detect issues with the construction of a formula and, at the user's discretion, automatically remove variables that would otherwise generate errors.  These quality checks include formatting variables with backticks, de-duplication, ensuring correspondence with the names of the variables in an associated data.frame, excluding categorical variables that would generate errors due to a lack of contrast or exceed a user-specified threshold for the maximum number of categories, and automatically removing interactions involving variables that should be excluded.  When directed by the user, these quality checks can be implemented to effectively reduce a formula to the subset of variables and interactions that would be appropriate for consideration in a statistical model.  The output of the function can be formatted as either a formula object or as a character.


```{r create.formula examples 1}
form.1 <- create.formula(outcome.name = outcome.name.awareness, input.names = NULL, input.patterns = NULL, dat = NULL)

form.1$formula
form.1$inclusion.table
form.1$interactions.table
```

```{r create.formula examples 2}
form.2 <- create.formula(outcome.name = outcome.name.awareness, input.names = ".", input.patterns = NULL, dat = snack.dat, include.intercept = FALSE)

form.2$formula
form.2$inclusion.table
form.2$interactions.table
```

```{r create.formula examples 3}
form.3 <- create.formula(outcome.name = outcome.name.awareness, input.names = input.names, input.patterns = NULL, dat = NULL, reduce = reduce, max.input.categories = max.input.categories, max.outcome.categories.to.search = max.outcome.categories.to.search, order.as = order.as, include.backtick = include.backtick, format.as = "character")

form.3$formula
form.3$inclusion.table
form.3$interactions.table
```


```{r create.formula examples 4}
form.4 <- create.formula(outcome.name = outcome.name.awareness, input.names = input.names, input.patterns = input.patterns, dat = snack.dat, reduce = reduce, max.input.categories = max.input.categories, max.outcome.categories.to.search = max.outcome.categories.to.search, order.as = order.as, include.backtick = include.backtick, format.as = format.as)

form.4$formula
form.4$inclusion.table
form.4$interactions.table
```

```{r create.formula examples 5}
form.5 <- create.formula(outcome.name = outcome.name.awareness, input.names = input.names, input.patterns = input.patterns, dat = snack.dat, reduce = TRUE, max.input.categories = max.input.categories, max.outcome.categories.to.search = max.outcome.categories.to.search, order.as = order.as, include.backtick = include.backtick, format.as = format.as)

form.5$formula
form.5$inclusion.table
form.5$interactions.table
```

```{r create.formula examples 6}
form.6 <- create.formula(outcome.name = outcome.name.satisfaction, input.names = input.names, input.patterns = input.patterns, dat = snack.dat, reduce = TRUE, max.input.categories = max.input.categories, max.outcome.categories.to.search = max.outcome.categories.to.search, order.as = order.as, include.backtick = include.backtick, format.as = format.as)

form.6$formula
form.6$inclusion.table
form.6$interactions.table
```

```{r create.formula.example 7}
form.7 <- create.formula(outcome.name = outcome.name.awareness, input.names = input.names, input.patterns = input.patterns, dat = snack.dat, interactions = interactions, force.main.effects = TRUE, reduce = TRUE, max.input.categories = max.input.categories, max.outcome.categories.to.search = max.outcome.categories.to.search, order.as = order.as, include.backtick = include.backtick, format.as = format.as, variables.to.exclude = "Gender")

form.7$formula
form.7$inclusion.table
form.7$interactions.table
```

```{r create.formula.example 8}
form.8 <- create.formula(outcome.name = outcome.name.awareness, input.names = NULL, input.patterns = NULL, dat = snack.dat, interactions = interactions, force.main.effects = TRUE, reduce = TRUE, max.input.categories = max.input.categories, max.outcome.categories.to.search = max.outcome.categories.to.search, order.as = order.as, include.backtick = include.backtick, format.as = format.as)

form.8$formula
form.8$inclusion.table
form.8$interactions.table
```

```{r create.formula.example 9}
form.9 <- create.formula(outcome.name = outcome.name.awareness, input.names = NULL, input.patterns = NULL, dat = snack.dat, interactions = interactions, force.main.effects = FALSE, reduce = TRUE, max.input.categories = max.input.categories, max.outcome.categories.to.search = max.outcome.categories.to.search, order.as = order.as, include.backtick = include.backtick, format.as = format.as)

form.9$formula
form.9$inclusion.table
form.9$interactions.table
```

```{r create.formula.example 10}

form.10 <- create.formula(outcome.name = outcome.name.awareness, input.names = NULL, input.patterns = NULL, dat = NULL, interactions = interactions, force.main.effects = FALSE, reduce = FALSE, max.input.categories = max.input.categories, max.outcome.categories.to.search = max.outcome.categories.to.search, order.as = order.as, include.backtick = include.backtick, format.as = format.as)

form.10$formula
form.10$inclusion.table
form.10$interactions.table
```


```{r create.formula.expample 11}
form.11 <- create.formula(outcome.name = outcome.name.awareness, input.names = c("Age", "Gender", "Income"), dat = snack.dat, variables.to.exclude = variables.to.exclude)

form.11$formula
form.11$inclusion.table
form.11$interactions.table
```

```{r create.formula.expample 12}
form.12 <- create.formula(outcome.name = outcome.name.awareness, input.names = input.names, dat = snack.dat, variables.to.exclude = c("Persona", "Region"), input.patterns = input.patterns, interactions = interactions, reduce = TRUE, force.main.effects = TRUE, include.backtick = "as.needed", format.as = 'formula')

form.12$formula
form.12$inclusion.table
form.12$interactions.table
```

Duplicates in input.names Example: if there is duplicates in input.name, for instance `r duplicated.inputs`, create.formula will deduplicate them and take only one varialbe.

```{r create.formula.example 12}
deduped.formula <- create.formula(outcome.name = outcome.name.awareness, input.names = duplicated.inputs, dat = snack.dat)

deduped.formula$formula
deduped.formula$inclusion.table
deduped.formula$interactions.table
```

Duplicates in interactions: if there are duplicates in interactions, for instance list(c("Age", "Income"), c("Age", "Income")), create.formula will deduplicate them and take only one interaction. 

```{r create.formula.example 13}
deduped.formula2 <- create.formula(outcome.name = outcome.name.awareness,input.names = duplicated.inputs2,dat = snack.dat,interactions = duplicated.interactions)

deduped.formula2$formula
deduped.formula2$inclusion.table
deduped.formula2$interactions.table
```

# Reduce Formula Example

```{r reduce.existing.formula examples 1}
reduce.form.1 <- reduce.existing.formula(the.initial.formula = form.3$formula, dat = snack.dat, format.as = "character")

reduce.form.1$formula
reduce.form.1$inclusion.table
reduce.form.1$interactions.table
```

```{r reduce.existing.formula examples 2}
reduce.form.2 <- reduce.existing.formula(the.initial.formula = form.4$formula, dat = snack.dat, format.as = "formula")

reduce.form.2$formula
reduce.form.2$inclusion.table
reduce.form.2$interactions.table
```
