---
title: "Introduction to formualic"
output: rmarkdown::html_vignette
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{Introduction to formualic}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(formulaic)
library(data.table)
library(knitr)
data("snack.dat", package = "formulaic")
```

```{r constant}
outcome.name.awareness <- "Awareness"
outcome.name.satisfaction <- "Satisfaction"
duplicated.inputs <- rep.int(x = "Age", times = 2)
duplicated.inputs2 <- c(rep.int(x = "Age", times = 2), "Income")
duplicated.interactions <- list(c("Age", "Income"), c("Age", "Income"))
input.names <- c("Age", "Gender", "Income", "Region", "Persona", "Typo")
input.patterns <- c("User", "BP_", outcome.name.awareness, outcome.name.satisfaction)
interactions <- list(c("Age", "Gender"), c("Age", "Income"), c("Age", "Gender", "Income"), c("Gender", "Inco"), c("Age", "Reg ion"))
reduce <- FALSE
max.input.categories = 20
max.outcome.categories.to.search = 4
order.as = "as.specified"
include.backtick = "as.needed"
format.as = "formula"
force.main.effects <- TRUE
```

## Data : snack.dat

The snack.dat was provided as an example on which to explore the basic functions of the formulaic package. Formatted as a data.table object, snack.dat contains 23000 observation and 23 columns. These data contain simulated information from a fictionalized marketing survey.  In this survey, a progression of questions are asked about the respondents' awareness, consideration, consumption, satisfaction with, and advocacy for different brands of snack foods.  Questions downstream of awareness, then consideration, and then consumption are only asked for those respondents who responded affirmatively to the previous question.  Otherwise the values are missing.  Brand Perception questions are rated on a scale from 0 to 10 and indicated with a name starting with the prefix BP.


```{r snack.dat}
dim(snack.dat)
names(snack.dat)
head(snack.dat)
```

## add.backtick Examples

It is a subsidiary function to create.fomula that add backticks to the variables. As a default, the include.backtick is set 'as.needed', which means that it only add backticks to the ones are required. User has the freedom to change the option to 'all'. Yet, it is only compatible when format.as != "formula".

As one can see, User ID, which requires backticks to be incorporated into a formula, is the only variable that is affected by the function when the included.backtick is set as 'as.needed', while every variable has backticks when it is set as 'all'. Note that this is only competible when 


```{r add.backtick}
formulaic:::add.backtick(x = names(snack.dat),include.backtick = 'all')
formulaic:::add.backtick(x = names(snack.dat),include.backtick = 'as.needed')
```

## create.formula examples

It automatically generates a formula based on the following parameters:
outcome.name, input.names, input.patterns, dat, interactions, force.main.effects, reduce, max.input.categories, max.outcome.categories.to.search, order.as, include.backtick, and format.as. 


```{r create.formula examples 1}
form.1 <- create.formula(outcome.name = outcome.name.awareness, input.names = NULL, input.patterns = NULL, dat = NULL)

form.1$formula
form.1$inclusion.table
form.1$interactions.table
```

```{r create.formula examples 2}
form.2 <- create.formula(outcome.name = outcome.name.awareness, input.names = ".", input.patterns = NULL, dat = snack.dat)

form.2$formula
form.2$inclusion.table
form.2$interactions.table
```

```{r create.formula examples 3}
form.3 <- create.formula(outcome.name = outcome.name.awareness, input.names = input.names, input.patterns = NULL, dat = NULL, reduce = reduce, max.input.categories = max.input.categories, max.outcome.categories.to.search = max.outcome.categories.to.search, order.as = order.as, include.backtick = include.backtick, format.as = "character")

form.3$formula
form.3$inclusion.table
form.3$interactions.table
```


```{r create.formula examples 4}
form.4 <- create.formula(outcome.name = outcome.name.awareness, input.names = input.names, input.patterns = input.patterns, dat = snack.dat, reduce = reduce, max.input.categories = max.input.categories, max.outcome.categories.to.search = max.outcome.categories.to.search, order.as = order.as, include.backtick = include.backtick, format.as = format.as)

form.4$formula
form.4$inclusion.table
form.4$interactions.table
```

```{r create.formula examples 5}
form.5 <- create.formula(outcome.name = outcome.name.awareness, input.names = input.names, input.patterns = input.patterns, dat = snack.dat, reduce = TRUE, max.input.categories = max.input.categories, max.outcome.categories.to.search = max.outcome.categories.to.search, order.as = order.as, include.backtick = include.backtick, format.as = format.as)

form.5$formula
form.5$inclusion.table
form.5$interactions.table
```

```{r create.formula examples 6}
form.6 <- create.formula(outcome.name = outcome.name.satisfaction, input.names = input.names, input.patterns = input.patterns, dat = snack.dat, reduce = TRUE, max.input.categories = max.input.categories, max.outcome.categories.to.search = max.outcome.categories.to.search, order.as = order.as, include.backtick = include.backtick, format.as = format.as)

form.6$formula
form.6$inclusion.table
form.6$interactions.table
```

```{r create.formula.example 7}
form.7 <- create.formula(outcome.name = outcome.name.awareness, input.names = input.names, input.patterns = input.patterns, dat = snack.dat, interactions = interactions, force.main.effects = TRUE, reduce = TRUE, max.input.categories = max.input.categories, max.outcome.categories.to.search = max.outcome.categories.to.search, order.as = order.as, include.backtick = include.backtick, format.as = format.as)

form.7$formula
form.7$inclusion.table
form.7$interactions.table
```

```{r create.formula.example 8}
form.8 <- create.formula(outcome.name = outcome.name.awareness, input.names = NULL, input.patterns = NULL, dat = snack.dat, interactions = interactions, force.main.effects = TRUE, reduce = TRUE, max.input.categories = max.input.categories, max.outcome.categories.to.search = max.outcome.categories.to.search, order.as = order.as, include.backtick = include.backtick, format.as = format.as)

form.8$formula
form.8$inclusion.table
form.8$interactions.table
```

```{r create.formula.example 9}
form.9 <- create.formula(outcome.name = outcome.name.awareness, input.names = NULL, input.patterns = NULL, dat = snack.dat, interactions = interactions, force.main.effects = FALSE, reduce = TRUE, max.input.categories = max.input.categories, max.outcome.categories.to.search = max.outcome.categories.to.search, order.as = order.as, include.backtick = include.backtick, format.as = format.as)

form.9$formula
form.9$inclusion.table
form.9$interactions.table
```

```{r create.formula.example 10}
form.10 <- create.formula(outcome.name = outcome.name.awareness, input.names = NULL, input.patterns = NULL, dat = NULL, interactions = interactions, force.main.effects = FALSE, reduce = FALSE, max.input.categories = max.input.categories, max.outcome.categories.to.search = max.outcome.categories.to.search, order.as = order.as, include.backtick = include.backtick, format.as = format.as)

form.10$formula
form.10$inclusion.table
form.10$interactions.table
```


Duplicates in input.names Example: if there is duplicates in input.name, for instance `r duplicated.inputs`, create.formula will deduplicate them and take only one varialbe.

```{r create.formula.example 11}
deduped.formula <- create.formula(
  outcome.name = outcome.name.awareness, 
  input.names = duplicated.inputs, 
  dat = snack.dat)
deduped.formula$formula
deduped.formula$inclusion.table
deduped.formula$interactions.table
```

Duplicates in interactions: if there is duplicates in interactions, for instance duplicated.interactions, create.formula will deduplicate them and take only one interaction. 

```{r create.formula.example 12}
deduped.formula2 <- create.formula(
  outcome.name = outcome.name.awareness,
  input.names = duplicated.inputs2,
  dat = snack.dat,
  interactions = duplicated.interactions)
deduped.formula2$formula
deduped.formula2$inclusion.table
deduped.formula2$interactions.table
```

# Reduce Formula Example

```{r reduce.existing.formula examples 1}
reduce.form.1 <- reduce.existing.formula(the.initial.formula = form.3$formula, dat = snack.dat, format.as = "character")
reduce.form.1$formula
reduce.form.1$inclusion.table
reduce.form.1$interactions.table
```

```{r reduce.existing.formula examples 2}
reduce.form.2 <- reduce.existing.formula(the.initial.formula = form.4$formula, dat = snack.dat, format.as = "formula")
reduce.form.2$formula
reduce.form.2$inclusion.table
reduce.form.2$interactions.table
```
